Notes and code examples to accompany "If the Null Fits, You Must Omit

### 2025-12-11 index.qmd
minor updates and cleanup of analysis; about to post docs

1.  **Calculate false positive/detection rates for the simulated null polarization
    model**, *mere agreement*—where polarization experiment participant opinions agree
    more following discussion than before. Recall this is modeled as a decrease in the
    standard deviation of latent opinions, but a constant mean latent opinion. Here, then, we calculate the false positive rate, $\alpha_e$ as a function of the effect-detection threshold, $d^*$, i.e., 
    $$
      \alpha_e(d^*) = \Pr\!\left( \hat{d} \ge d^{*} \mid \mathcal{M}_{0} \right).
    $$
    In the study, we calculate the right-hand side as the frequency of the 1000 trials where $\hat{d} \ge d^*$ for simulations of experiment, $e$. 
2.  **Find an effect–detection threshold**, $d^*(\alpha)$, such that, under the null
    model, the probability of observing an estimated effect size exceeding
    \$d\^\*(\\alpha)\$ is some target false positive rate, $\alpha^* = 0.05$, for example: 
    $$
     d^* = \min~d,~~\text{subject to}~\Pr\!\left( \hat{d} \ge d \mid \mathcal{M}_{0} \right)  \le\alpha^*.
     $$ 
     This yields a null-model–calibrated false–positive rate for detecting a group
    polarization effect.

@@

In the second step of our generative model analysis I calculate

>>>>

1.  **Calculate false positive/detection rates for the simulated null polarization
    model**, *mere agreement*—where polarization experiment participant opinions agree
    more following discussion than before. Recall this is modeled as a decrease in the
    standard deviation of latent opinions, but a constant mean latent opinion. Here, then, we calculate the false positive rate, $\alpha_e$ as a function of the effect-detection threshold, $d^*$, i.e., 
    $$
      \alpha_e(d^*) = \Pr\!\left( \hat{d} \ge d^{*} \mid \mathcal{M}_{0} \right).
    $$ {#eq-false-positive-rate}
    In the study, we calculate the right-hand side as the frequency of the 1000 trials where $\hat{d} \ge d^*$ for simulations of experiment, $e$. 
2.  **Find an effect–detection threshold**, $d^*(\alpha)$, such that, under the null
    model, the probability of observing an estimated effect size exceeding
    \$d\^\*(\\alpha)\$ is some target false positive rate, $\alpha^* = 0.05$, for example: 
    $$
     d^* = \min~d,~~\text{subject to}~\Pr\!\left( \hat{d} \ge d \mid \mathcal{M}_{0} \right)  \le\alpha^*.
     $$ {#eq-threshold-spec}
     This yields a null-model–calibrated false–positive rate for detecting a group
    polarization effect.

@@

In the second step of our generative model analysis I calculate the solution to
the problem in @eq-threshold-spec, finding the effect-detection threshold, $d^*$, that achieves a target false positive rate, $\alpha^* = 0.05$.
